---
output: html_document
editor_options: 
  chunk_output_type: inline
---

# - Creation of the package structure

## Initiation the global architecture

Now you have an working environment correctly configured and we have taken a few moment to thing about our package aim. We can go deeper and initiate the structure of our new package from an active R session. You don't have to take care about the configuration of your local session because the function associated will be create a clean new project for your package. First, load the [devtools](https://devtools.r-lib.org/){.external target="_blank"} and by association is friend [usethis](https://usethis.r-lib.org/){.external target="_blank"}.

```{r}
#| label: load-devtools-and-usethis
#| echo: true
library(devtools)
```

Now launch the following command to initiate the global architecture of your package. For the training, my new R package will have the acronym "macfly" for "My Anwesone paCkage For Learning with Yearning".

```{r}
#| label: create-package
#| echo: true
create_package(path = "~/macfly")
```

Another solution with RStudio is to use the interface through the *file* menu, *New Project...*, *New Directory*, *R Package*, file the field "Package name:", click on "Open in new session" and finish with \*Create Project". Even is the approach is similar, the click-button interface create in addition a template function with a proposal for the documentation associated. In opposite, the command line creation include by default a .gitignore file (for Git use). It's not true important but it could be a bit better to use the command line because we will learn how to create a R function template later and in your developer life it could be very rare (and dangerous) when you don't use a Git associated to your work.

By default, RStudio open a fresh session associated to the R environment of your package and you should need to call again `library(devtools)`.

::: callout-note
For your information, you will find in the [usethis](https://usethis.r-lib.org/index.html){.external target="_blank"} package a function [`create_tidy_package()`](https://usethis.r-lib.org/reference/tidyverse.html){.external target="_blank"} that extends `create_package()` by applying many of the tidyverse development conventions as possible. I don't recommend using that way, instead of you know exactly what you do and to save time in the future, because it's better to really understand each step of the package creation.
:::

::: callout-important
At least, you could see sometime a function call `package.skeleton` from the [utils](https://www.rdocumentation.org/packages/utils/versions/3.6.2/topics/package.skeleton){.external target="_blank"} R package. Even if this function create a package structure it's an "old way" to do and you should have some trouble if you use it, especially with you want to use the tidyverse philosophy.
:::